Tolerant  to ethnic diversity, intolerant to intolerance

### 2020-01-23 ethnic_value_rum.Rmd
Update ethnic_value_rum.Rmd
After meeting AF Groningen 13

---
output:
  bookdown::pdf_document2:
    toc: false
 #   keep_tex: false
    fig_caption: yes
title: "Preferences interact: what consequences for a Schelling scenario?"
# subtitle: "Target: BIGSSS Research Day"
author:  Rocco Paolillo, Andreas Flache, Jan Lorenz
# date:  "`r  as.Date(Sys.time(), '%d %B %Y')`"
# "`r  as.Date(Sys.time(), '%d %B %Y')`"
language: en-GB
bibliography: "references.bib"
always_allow_html: yes
header-includes:
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{xcolor} 
- \usepackage{amsmath}
- \usepackage{graphicx}
---

@@

@@

* Schelling's model and previous work. Interaction of preferences.
Schelling's model
Literature update
@paolillo2018 Diverse dimensions can define individuals and their similarity. Results from @paolillo2018
@hedstrom2009oxford, chapter 11, 12

* Why an additional characteristic than ethnicity, literature in residential studies

* Update: justification for random utility models, literature in ABM
Better define the individual  decision process
Link with empirical data
@huang2014review
@klabunde2016decision
@bruch2019choice

* Research question
How would the interaction between $\beta$ for different dimensions affect Schelling's scenarios

# Model and Experiments

@@

* Calculation utility and probability function

To explore what levels of $\beta$ should not be excluded from the manipulations, I built a specific model with a unique beta for ethnic composition ($\beta_{ie}$ ) that does not change among agents. At each run I incremented degree of $\beta_{ie} \in[0,100]$ and let run for 1000 ticks with agents asynchronously relocating (not 1 agent for 1 tick, all to complete the procedure to restart). In each condition linear function was used: the preference peak of agents was $100\%$ for ethnically similar

To explore what levels of $\beta$ should not be excluded from the manipulations, I made a specific model "rum_eth.nlogo" (in material), with a unique $\beta$ for all agents. I run the simulation for 1000 ticks (not 1 agent 1 tick):
* linear function: peak preference for all agents at $100 \%$
* density == $70 \%$
* equal distribution of value orientation in each ethnic group
* changing $beta_ie$ at each run $beta_ie \in[0,100]$

```{r one-beta, fig.cap="Comparison of ethnic segregation for one beta-ethnic, not interaction. Used one-beta.csv from simulation rum-eth.nlogo" , include=TRUE}

a <- read.csv("one-beta.csv",sep = ",",skip = 6)
names(a)[names(a) == "mean..count..turtles.on.neighbors..with..ethnicity....ethnicity..of.myself....count..turtles.on.neighbors....of.turtles.with..count..turtles.on.neighbors.....1." ] <- "eth_seg"

a %>% filter(ticks == 0  | ticks == 500  | ticks == 100 | ticks == 1000)  %>%  filter(beta.ie >= 0 & beta.ie <= 15) %>% ggplot(aes(x = beta.ie, y = eth_seg)) + geom_bar(stat="identity") + facet_wrap(~ ticks, labeller = label_both)

```{r att_income , include=TRUE}

f_att <- function(i,b){   # utility function as single-peaked
 A <- 1 / (1 + exp(-b(i)))
}

  b <- 0.5

a_0 <- f_cst(a,0)
a_1 <- f_cst(a,0.1)
a_2 <- f_cst(a,0.2)
a_3 <- f_cst(a,0.3)
a_4 <- f_cst(a,0.4)
a_5 <- f_cst(a,0.5)
a_6 <- f_cst(a,0.6)
a_7 <- f_cst(a,0.7)
a_8 <- f_cst(a,0.8)
a_9 <- f_cst(a,0.9)
a_10 <- f_cst(a,1)

Prob_A <- c(a_0,a_1,a_2,a_3,a_4,a_5,a_6,a_7,a_8,a_9,a_10)
income <- seq(0,i, by = 0.1)

plot(price,na.omit(utility),ylab = "utility_ec",xlab="price", main = paste0("income =",i))

```{r income_price , include=TRUE}

# tentative of utility economic, given by the difference (income - price), so to represent how convenient a location is. 
# ideally identity equation: if income < price, probability would be 0. If beta is negative, the probability could still be computed, which is nonsense

f_cst <- function(i,p){   # utility function as single-peaked
 if (i < p){
    U <- NA}else{
  U <- ((i - p) / i)}
}

  i <- 0.5

i_0 <- f_cst(i,0)
i_1 <- f_cst(i,0.1)
i_2 <- f_cst(i,0.2)
i_3 <- f_cst(i,0.3)
i_4 <- f_cst(i,0.4)
i_5 <- f_cst(i,0.5)
i_6 <- f_cst(i,0.6)
i_7 <- f_cst(i,0.7)
i_8 <- f_cst(i,0.8)
i_9 <- f_cst(i,0.9)
i_10 <- f_cst(i,1)

utility <- c(i_0,i_1,i_2,i_3,i_4,i_5,i_6,i_7,i_8,i_9,i_10)
price <- seq(0,i, by = 0.1)

plot(price,na.omit(utility),ylab = "utility_ec",xlab="price", main = paste0("income =",i))

>>>>

---
output:
  bookdown::pdf_document2:
    toc: false
 #   keep_tex: false
    fig_caption: yes
title: "Tolerant  to ethnic diversity, intolerant to intolerance"
subtitle: "Title to define"
author:  Rocco Paolillo, Andreas Flache
# date:  "`r  as.Date(Sys.time(), '%d %B %Y')`"
# "`r  as.Date(Sys.time(), '%d %B %Y')`"
language: en-GB
bibliography: "references.bib"
always_allow_html: yes
header-includes:
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{xcolor} 
- \usepackage{amsmath}
- \usepackage{graphicx}
---

@@

# INTRO FOCUSED ON SCHELLING, MUST BE ON HYBRID SEGREGATION

@@

STRESS A LINEARITY APPROACH vs SYSTEMIC APPROACH. Residential segregation can be defined as the unequal and non random distribution of people  belonging to different groups within the spatial units of a city [@bruch2009segregation;@clark2008understanding]. The interest of the social sciences on the topic is due the consequences it can have on the chance of interaction and integration between groups [@bolt2010linking], with exclusion of minorities in deprived neighborhoods and perpetuation of inequalities through generations [@van2012understanding]. A key research question in literature is on the dynamics of emergence of residential segregation. Two hystorical hypothesis are the spatial assimilation theory and the place stratification theory. According to the former, spatial distance reflects the social distance between groups, and can change as effect of social moblity (@coulter2019ethnic), while the latter focuses on the strategies of gatekeepers, e.g. sellers, to prevent undesired newcomers to neighborhood (@horr2018ethnic). Both have in common the assumption that residential  segregation is the intended result of individual behavior. A third perspective can be that of social complexity and Schelling's agent-based model of residential segregation, which addresses residential segregation as a stable property of the aggregated behavior of people who seek to satisfy even mild homophily preferences within spatial constraints [@schelling1969;@schelling1971;@schelling1978]. The paper here presented follows this stream of literature and aims at contributing to Schelling's type dynamics. \par

Diversity in cities: how to navigate the difference: Heterogeneity in ethnic preferences and competing preferences

Lead Question (ex):
* how can we explain hybrid ethnic segregation patterns? (The Economist)
* hybrid segregation on value and ethnic for different types of people

Assumption: 
* people differ in their degree of ethnic preferences (lit. @xie2012modeling, @hatna2015combining)
* peole have different preferences and sort out of different factors [@bruch2009segregation]

@xie2012modeling: use different threshold size according to class from Guttman, the 6th class not conforming to Guttman linear function and rum. Their concept of heterogeneity is on the different threshold levels (within 5 classes, with the 5th being tolerant to $57 \%$ to $100 \%$) black, so indifferent), with the consequence to decrease segregation. This trend holds also when continuous preferences are included, meaning cumulative threshold between agents. We would modify $\beta$, decomposing a random utility model, assuming the desired concentration peak 1 and varying $\beta$ \par

In  the definition of competing preferences, stress that the value-orientation across ethnicity is defiant to Schelling's dichotomous definition of similarity (see @clark1991residential for a critical view)

@clark1991residential old: for historical excursus
@chen2005emergence: good perspective to position with value similarity (not in their paper). Update literature, consider heterogeneity within the same race

Define need of value similarity

Role of relative group size for both the ethnic and the value dimension

```{r model, fig.pos ='h!',out.width='50%',out.height='90%',fig.show='hold',fig.align='center', fig.cap= "Draft of agents attribution and emerging segregations, to revise.",  echo = FALSE, include=TRUE}
knitr::include_graphics("C:/Users/rocpa/OneDrive/Desktop/AF/ethnic-value_multinomial/material/model_pr.jpg")
```

# Model

@@

$\beta^\Box_e \geq \beta^\circ_e$

# Experiments

Study 1: heterogeneity of ethnic preferences in agents \par
* Tune $\beta_e$ for all agents, with condition $\beta^\Box_e \geq \beta^\circ_e$
* Same conditions, increase $\beta_v$

The concept of heterogeneity is reflected into the experiments: \par
* individual: $\beta^\Box_e \geq \beta^\Box_v$, $\beta^\circ_v \geq \beta^\circ_e$  \par
* within ethnic group:  $\beta^\Box_e \geq \beta^\circ_e$, $\beta^\circ_v \geq \beta^\Box_v$, should specify the ethnic group  \par
NO * between ethnic group: $\beta^{\Box B}_e \geq \beta^{\Box R}_e$ (and implying $\beta^{\Box B}_e \geq \beta^{\circ R}_e$  ) NO \par

These are the conditions kept in the experiments, ranging $\beta \in{0,10}$  \par

Think about it.
